---
title: "Wrangle Spring 2023 Data"
output:
  rmarkdown::html_vignette:
    css:
      - !expr diffobj::diffobj_css()
      - !expr system.file("rmarkdown", "templates", "html_vignette", "resources", "vignette.css", package = "rmarkdown")
---vignette: >
  %\VignetteIndexEntry{Wrangle Spring 2023 Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

```{r setup}
library(wihera)
.TEMPLATE_FILE <- system.file("extdata",
                              "2023 HERA Data Submission form.xlsx",
                              package = "wihera")

.WI_STATE_FILE <- system.file("extdata",
                              "2023 WI State U A&M HERA Data Submission form.xlsx",
                              package = "wihera")

```

Here are the data files that I will use.

```{r import_the_data}
template_sheets <- read_template(.TEMPLATE_FILE)

wi_state_sheets <- read_template(.WI_STATE_FILE)
```

## Things Tableau Didn't Catch:

- places where numerators were excluded (because < 10) but denominator

## Template Checking

There is a template sheet that each school was supposed to use to submit its data.
Like many spreadsheets, it encodes some of its information in cell locations.
If the schools changed the order of any of the column or row labels then there will be a mismatch between what they intended to report and what the automated import system will say that they reported.
The following code checks each submission against the template to catch mismatches.

```{r template_checking, results="asis"}
sheet_checks <- TEMPLATE_REGIONS$`2023` |>
    purrr::imap( ~ check_school(wi_state_sheets[[.y]],
                                .x,
                                template_sheets[[.y]])
    )

purrr::iwalk(
    sheet_checks,
    function(.check, .sheet) {
        cat("\n\n#### ", .sheet, "\n\n", sep = "")
        purrr::walk(.check$messages,
                    ~ purrr::iwalk(., mismatch_as_markdown, "#####"))
    }
)
```

## Definitions from the Template

These definitions are intended as guidance for completing the template.
If alternative definitions or measures are more appropriate for your institution, please meet the intent as best you can and include notes of differences.

### Populations

Population definitions are intended to align with IPEDS where applicable.

#### First-time cohort

First-time degree/certificate-seeking undergraduates entering the fall term.
Students should have no previous college credits other than credits earned while enrolled in high school.
Students admitted for fall who begin enrollment in the summer also may be included.
If the first time cohort is less than 10, suppress the data.

#### Full-time

Enrolled for 12 or more semester credits (or 12 or more quarter credits or 24 or more contact hours a week) at the beginning of the term.

#### Part-time

Enrolled for less than 12 semester credits (or less than 12 quarter credits or less than 24 contact hours a week) at the beginning of the term.

```{r enrollment_status_factor}
ENROLLMENT_STATUS_LEVELS <- c(
    "Full-time",
    "Part-time"
)
```

#### Race/ethnicity

As reported to IPEDS

```{r race_factor}
RACE_LEVELS <- c(
    "Hispanic",
    "Black, non-Hispanic",
    "White, non-Hispanic",
    "Asian",
    "Native Hawaiian or other Pacific Islander",
    "American Indian/Alaska Native",
    "Two or More Races",
    "Unknown",
    "Non-resident Alien"
)

```

#### Gender

As reported to IPEDS

```{r gender_levels}
GENDER_LEVELS <- c(
    "Male",
    "Female"
)
```

#### Age

As of the beginning of the term of entry.
Students under age 17 may be included in the Age: 17 to 19 category.

```{r age_levels}
AGE_LEVELS <- tibble::tribble(
    ~ Level,           ~ `Lower Age`, ~ `Upper Age`,
    "17-19 years old",            0L,           19L,
    "20-24 years old",           20L,           24L,
    "Age 25 and over",           25L,          999L,
    "Unknown",                    NA,            NA
)
```

#### Income

Received Pell Grant in the term of entry.
Eligibility for Pell Grant in the term of entry may be used as an alternative.

### Degree/certificate programs

Program definitions are intended to align with IPEDS where applicable.

#### Degree/certificate program of less than 2 years

As defined in IPEDS Graduation Rate Survey

#### Degree/certificate program of at least 2 years but less than 4 years

As defined in IPEDS Graduation Rate Survey

#### Bachelor's degree

Bachelor's or equivalent degree as defined in IPEDS Graduation Rate Survey

## Measures

Measures are intended to align with IPEDS where applicable.

### Retention

#### Retention

Number of students in the cohort who enrolled at the beginning of the next fall term.
Include students who graduated/completed their program within one year, if any. 

```{r retention_schema}
retention_schema <- list(
    sheet = "Retention",
    range = "A6:C41",
    col_names = c("Population",
                  "First-time Cohort",
                  "Enrolled following fall term"),
    col_types = c("text", "numeric", "numeric"),
    na = c("", "NA", "N/A", "DS")
)
```

### Momentum in First Year

#### Within first term

Within the term of entry, e.g., fall 2017

#### Within first year

For students entering fall 2017, includes summer 2017 (if enrolled), fall 2017, winter 2017-18 (if offered), spring 2018, and summer 2018.

#### Completed

Passing or earning a credit that counts toward credit requirements for graduation, as a result of enrolling in a course at the institution.
Students earning credits through placement or prior learning should be included in the Exempt category, not here.
Remedial/developmental courses that do not contribute credit hours toward degree requirements should not be included.

#### Exempt

Students who do not need to take the course either because their program does not require it or as a result of placement into a higher level course, AP credit, dual credit earned while enrolled in high school, CLEP, IB, or other reasons (please note). 

#### Credits

Semester credit hours.
Institutions using quarter hours should divide quarter hours by 1.5 to convert to semester hours.

#### Gateway course

First credit bearing college course that applies to a requirement for a certificate or degree.
Credit for the course counts toward credit requirements for graduation.
The course has no credit bearing prerequisites, although it may have remedial / developmental prerequisites.

#### Gateway math course

A gateway course that satisfies a mathematics competency or quantitative literacy/reasoning requirement for graduation.

#### Gateway English course

A gateway course that satisfies an English or similar requirement for graduation.

#### Declared major/program

Students admitted directly to a program, or students who complete any paperwork required by the institution to declare a major or program.

```{r momentum_schema}
momentum_schema <- list(
    sheet = "Momentum in First Year",
    range = "A6:J41",
    col_names = c("Population",
                  "First-time cohort \n(Fall 2021)",
                  "Completed credit hours within first term \n(15+ if entered full-time, 9+ if entered part-time)",
                  "Completed credit hours within first year \n(30+ if entered full-time, 15+ if entered part-time)",
                  "Completed gateway math course within first year",
                  "Exempt from gateway math course",
                  "Completed gateway English course within first year",
                  "Exempt from gateway English course",
                  "Completed or exempt from gateway math AND completed or exempt from gateway English AND completed credit hours within first year (30+ if entered full-time, 15+ if entered part-time)",
                  "Declared a major/program within first year" ),
    col_types = c("text", rep("numeric", 9)),
    na = c("NA", "N/A", "DS")
)
```

### Math Pathways

#### Exempt

See above.

#### College Algebra

A gateway math course that provides the algebra skills required for calculus.

#### Quantitative Reasoning

A gateway math course that emphasizes mathematical reasoning and its practical application through solving quantitative problems in a wide array of everyday life situations and making and evaluating logical arguments supported by quantitative evidence.

#### Statistics

A gateway math course including descriptive statistics, elementary probability, general and sampling distributions, and the fundamentals of statistical inference, including confidence intervals and hypothesis testing, simple regression, and correlation.

#### Technical Math

(If you have a course of this type, please provide a definition.)

#### Other

(For example, math for elementary education majors.
Insert additional columns as needed and explain in the notes.)

#### More than one gateway math course

This column is provided to allow the sum of students by math pathways to equal the total who completed a math gateway course that is reported in Momentum in First Year.

```{r math_pathways_schema}
math_pathways_schema <- list(
    sheet = "Math Pathways",
    range = "A8:I11",
    col_names = c("Population",
                  "First-time cohort \n(Fall 2021)",
                  "Exempt",
                  "College Algebra",
                  "Quantitative Reasoning",
                  "Statistics",
                  "Technical Math",
                  "Other Course (total)",
                  "More than one gateway math course"
    ),
    col_types = c("text", rep("numeric", 8)),
    na = c("NA", "N/A", "DS")
)


math_pathways_offered_schema <- list(
    sheet = "Math Pathways",
    range = "C7:I7",
    col_names = c("Exempt",
                  "College Algebra",
                  "Quantitative Reasoning",
                  "Statistics",
                  "Technical Math",
                  "Other Course (total)",
                  "More than one gateway math course"
    ),
    col_types = "text",
    na = c("NA", "N/A", "DS")
)
```

### Graduation

#### Graduation

Number of students in the cohort who earn the degree/certificate within 100% and 150% of the expected (full-time) degree/certificate length.
Note that the count for 150% is cumulative, that is, it includes students completing in 100% time.

### Graduates' Time and Credits	

#### Average credits to degree

Start with the degrees/certificates awarded in a specified year and determine the total number of semester credit hours each student completed since first enrolling.
Include first-time degree/certificate-seeking undergraduates entering in any term; exclude students entering as transfer students.
Average the number of semester credit hours across students and report by degree type.

#### Average time to degree

Start with the degrees/certificates awarded in a specified year and determine the total number of years and months elapsed from the first date of entry to the date of completion.
Partial years should be expressed as a decimal.
Include first-time degree/certificate-seeking undergraduates entering in any term; exclude students entering as transfer students.
Average the number of years across students and report by degree type.

```
