---
title: "Wrangle Spring 2023 Data"
output:
  rmarkdown::html_vignette
---vignette: >
  %\VignetteIndexEntry{Wrangle Spring 2023 Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
.TEMPLATE_FILE <- system.file("extdata",
                              "2023_HERA_Data_Submission_form.xlsx",
                              package = "wihera")

.WI_STATE_FILE <- system.file("extdata",
                              "2023_Example_HERA_Data_Submission_form.xlsx",
                              package = "wihera")

```

The first step is to read in the template and the example file as all-text matrices.

```{r import_the_templates}
template_sheets <- wihera::read_template(.TEMPLATE_FILE)

wi_state_sheets <- wihera::read_template(.WI_STATE_FILE)
```

## Things Tableau Didn't Catch:

- places where numerators were excluded (because < 10) but denominator

## Template Checking

There is a template sheet that each school was supposed to use to submit its data.
Like many spreadsheets, it encodes some of its information in cell locations.
If the schools changed the order of any of the column or row labels then there will be a mismatch between what they intended to report and what the automated import system will say that they reported.
The following code checks each submission against the template to catch mismatches.

```{r template_checking, results="asis"}
sheet_checks <- wihera::TEMPLATE_REGIONS_2023 |>
    purrr::imap( ~ wihera::check_school(wi_state_sheets[[.y]],
                                        .x,
                                        template_sheets[[.y]])
    )

purrr::iwalk(
    sheet_checks,
    function(.check, .sheet) {
        cat("\n\n#### ", .sheet, "\n\n", sep = "")
        purrr::walk(.check$messages,
                    purrr::iwalk,
                    wihera::mismatch_as_markdown,
                    "#####")
        cat("\n")
    }
)
```

Note that the main source of mismatches is the type of line-ending character.
To be clear, I created the "Example" file using Excel's "Save As..." command.
Microsoft is **THE WORST!**

## Import the Data

The next step is to use each of the sheet-specific import functions to read the data.

```{r import_the_data}
example_data <- list(
    "Graduates' Time and Credits" = wihera::read_grads_times_credits,
    "Graduation - Bachelor's Degrees" = wihera::read_graduation_bachelors,
    "Graduation - Programs < 4 years" = wihera::read_graduation_cc_or_tech,
    "Math Pathways" = wihera::read_math_pathways,
    "Momentum in First Year" = wihera::read_momentum,
    "Retention" = wihera::read_retention
) |>
    purrr::map(~ .(.WI_STATE_FILE))
```

once imported, we can convert them into the format that Tableau will expect

```{r prepare_the_data_for_tableau}
dashboard_data <- list(
    "Graduates' Time and Credits" = wihera::dashboardify_grads_times_credits,
    "Graduation - Bachelor's Degrees" = wihera::dashboardify_graduation_bachelors,
    "Graduation - Programs < 4 years" = wihera::dashboardify_graduation_cc_or_tech,
    "Math Pathways" = wihera::dashboardify_math_pathways,
    "Momentum in First Year" = wihera::dashboardify_momentum,
    "Retention" = wihera::dashboardify_retention
) |>
    purrr::imap(~ .x(example_data[[.y]])) |>
    purrr::list_rbind(names_to = "Source Tab") |>
    wihera::relocate_for_dashboard()
```
